Binning Strategies and Related Loss for Large Data

### 2015-05-06 Paper.Rnw
frequency loss definition

\begin{itemize}
\item {\it Spatial Loss}, $L^\text{S}$, occurs when points $(x_i, y_i)$ for observations $i \in \{1,\dots,n\}$ in the data set are reduced to a set of tiles centered at $(x_\ell^\ast, y_\ell^\ast)$ for bins $\ell \in \{1,\dots,\mathscr{L}\}$. By displaying frequency information using shaded tiles instead of individual points there is a loss of information about the exact location of the points. %\hh{XXX the frequency information is not what causes the loss}
\ktm{
\item {\it Frequency Loss}, $L^\text{F}$, occurs when bin counts, $c_\ell \in \{1,\dots,\mathscr{L}\}$ are not mapped to a continuous shading scale. While shade can be \textit{rendered} continuously in HSV color space, thus representing frequency exactly, this information is extracted via imperfect human perception. A discrete color scale may be used to aid the ability to read a binned scatterplot, however this is effectively a second stage of binning where shade displays binned frequencies, $b_C(c_\ell)$, for spatial bins $\ell \in \{1,\dots,\mathscr{L}\}$, thus losing precision in the display of the true frequencies.
}
\end{itemize}

\hh{XXX ok, I don't quite agree with the description of frequency loss here. (a) the rendering on a continuous color scheme IS precise. but on a cognitive level we are not able to read the rendered values at the same precision. This effectively intorduces a second stage of binning ... then we are on the same page again.} \ktm{Is this new description any better?}

>>>>

\begin{itemize}
\item {\it Spatial Loss}, $L^\text{S}$, occurs when points $(x_i, y_i)$ for observations $i \in \{1,\dots,n\}$ in the data set are reduced to a set of tiles centered at $(x_\ell^\ast, y_\ell^\ast)$ for bins $\ell \in \{1,\dots,\mathscr{L}\}$. By displaying frequency information using shaded tiles instead of individual points there is a loss of information about the exact location of the points. %\hh{XXX the frequency information is not what causes the loss}
\ktm{
\item {\it Frequency Loss}, $L^\text{F}$, occurs when bin counts, $c_\ell \in \{1,\dots,\mathscr{L}\}$ are not mapped to a continuous shading scale. While shade can be \textit{rendered} continuously in HSV color space, thus representing frequency exactly, a human reader can not \textit{extract} this information at the same precision due to limitations of  human cognition.  %A discrete color scale may be used to aid the ability to read a binned scatterplot, however this is effectively 
In order to model these limitations we 
introduce a second stage of binning by using a discrete color scale for displaying binned frequencies, $b_C(c_\ell)$, $\ell \in \{1,\dots,\mathscr{L}\}$. %This loss in precision is then quantified as {\it Frequency loss}. %, thus losing precision in the display of the true frequencies.
}
\end{itemize}

%\hh{XXX ok, I don't quite agree with the description of frequency loss here. (a) the rendering on a continuous color scheme IS precise. but on a cognitive level we are not able to read the rendered values at the same precision. This effectively intorduces a second stage of binning ... then we are on the same page again.} \ktm{Is this new description any better?}
